---
title: "Detail - Z Scaling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detail - Z Scaling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.width  = 8,
  fig.height = 6
)
library(ggplot2)
library(rgl)
library(ggrgl)
```


Z scaling
-----------------------------------------------------------------------------

Variables may be mapped to the `z` direction just like the `x` and `y` aesthetics
in `ggplot2`.

The behaviour is a little different though in that:

* `x` and `y` aesthetics are naturally confined to the size of the page.
* The `z` direction has no natural boundary in a plot


Default z-position scales: `scale_z_continuous()` and `scale_z_discrete()`
-----------------------------------------------------------------------------

By default z values are scaled into the range [1, 100].  

How high something at
the maximum z coordinate actually appears on a given plot depends 3 things.

1. The `range` argument to `scale_z_continuous()`/`scale_z_discrete()`
2. the `zscale` argument to `devoutrgl::rgldev()` - which defaults to 1, and further
   applies a multiplicative scale factor to the extents in the z-direction.
3. the `dpi` argumnet to `devoutrgl::rgldev()`. This defaults to `dpi=72` which
is useful for screen presentation.  Higher values, which may be useful for 
saving to PNG, will cause a distortion in the Z scaling.


Default Z Range: `range = c(1, 100)`
-----------------------------------------------------------------------------

```{r}
n <- seq(1, 1000, length.out = 20)
plot_df <- data.frame(x=n, y=n, z=n+1000)

p <- ggplot(plot_df) + 
  geom_point_z(aes(x=x, y=y, z=z, colour = z), extrude = TRUE, extrude_face_alpha = 0.25)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Render Plot in 3d with {devoutrgl}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
devoutrgl::rgldev(fov = 30, view_angle = -30)
p
invisible(dev.off())
```

**Use your mouse, mouse buttons and scrollwheel to manipulate the 3d plot**

```{r echo=FALSE}
if (!interactive()) rgl::rglwidget()
```


Custom Z Range: `range = c(1, 500)`
-----------------------------------------------------------------------------

```{r}
p <- ggplot(plot_df) + 
  geom_point_z(aes(x=x, y=y, z=z, colour = z), extrude = TRUE, extrude_face_alpha = 0.25) + 
  scale_z_continuous(range = c(1, 500))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Render Plot in 3d with {devoutrgl}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
devoutrgl::rgldev(fov = 30, view_angle = -30)
p
invisible(dev.off())
```

**Use your mouse, mouse buttons and scrollwheel to manipulate the 3d plot**

```{r echo=FALSE}
if (!interactive()) rgl::rglwidget()
```



Custom Z Range: `range = c(100, 300)`
-----------------------------------------------------------------------------

```{r}
p <- ggplot(plot_df) + 
  geom_point_z(aes(x=x, y=y, z=z, colour = z), 
               extrude = TRUE, extrude_face_alpha = 0.25, extrude_z=100) + 
  scale_z_continuous(range = c(100, 200))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Render Plot in 3d with {devoutrgl}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
devoutrgl::rgldev(fov = 30, view_angle = -30)
p
invisible(dev.off())
```

**Use your mouse, mouse buttons and scrollwheel to manipulate the 3d plot**

```{r echo=FALSE}
if (!interactive()) rgl::rglwidget()
```





Default Z Range with `xscale > 1`: `range = c(1, 100)`
-----------------------------------------------------------------------------

Set `rgldev(zscale = ...)` to change the z scaling at the device level.

```{r}
n <- seq(1, 1000, length.out = 20)
plot_df <- data.frame(x=n, y=n, z=n+1000)

p <- ggplot(plot_df) + 
  geom_point_z(aes(x=x, y=y, z=z, colour = z), extrude = TRUE, extrude_face_alpha = 0.25)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Render Plot in 3d with {devoutrgl}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
devoutrgl::rgldev(fov = 30, view_angle = -30, zscale = 10)
p
invisible(dev.off())
```

**Use your mouse, mouse buttons and scrollwheel to manipulate the 3d plot**

```{r echo=FALSE}
if (!interactive()) rgl::rglwidget()
```
